### 2014-02-05 paper/appendix.tex
paper start

\newpage
\appendix
\section{Injection Measurements}
We injected the copper a distance of 1.16m (according to the spiral length calculator at http://www.giangrandi.ch/soft/spiral/spiral.shtml ) in a spiral whose cross-section was a square of area $9mm^2$.  I suspect that is about as far as we can go without using a vacuum at the end.

\section{Objet 260 Connex Digital Materials}
Softer materials (concentration of > 65% Tango-series material) do not easily accept the copper paint (it cracks when bent and is easy to wash off).  This may preclude flex sensors made of interior copper paint (conductive thread or other materials may be able to be used for some parts of this).

\section{Material Properties}

\begin{table*}[t]
\begin{tabular}{ l || l | p{3cm} | l | p{4cm} }
  \hline
  Name & Material & Resistance & Drying Time & Application Notes\\ \hline
  CuPro-Cote Coating & Copper & 2$\Omega$/inch & O(1 day) & Syringe \\ \hline
  Spectra 360 Electrode Gel & Liquid/Electrolytes & 125k$\Omega$/inch & O(hours) & Does not conduct dry \\ \hline
  Wire Glue & Carbon/Graphite &  23.6k$\Omega$/inch & O(minutes) & Syringe, very runny \\ \hline
  Bare Conductive Electric Paint & Carbon/Graphite & 110$\Omega$/inch & O(days) & Syringe \\ \hline
  Homemade Conductive Paint & Carbon/Graphite & 120$\Omega$/inch & O(hours) & \parbox{4cm}{Too thick for syringe,\\ apply externally with brush} \\ \hline
  Conductive Thread & Steel & \parbox{3cm}{1.8$\Omega$/inch taut \\ 2.5$\Omega$/inch loose} & N/A & \parbox{4cm}{Difficult to feed through\\turns} \\ \hline
  Solder Paste & Lead & 2$\Omega$/inch & N/A & \parbox{4cm}{Too thick for syringe,\\ must bake to conduct} \\ \hline
\end{tabular}
\end{table*}