### 2014-06-29 paper/conclusion.tex
making changes thanks to reviews

\section{Conclusion}

Pipe mediation opens new interactive device prototyping opportunities both for makers with consumer-grade 3D printers and research and industrial labs with higher-end machines.  While today's 3D printers are not yet able to fabricate active components in-place, we suggest that 3D printed interactive devices can be created with redirection of active input and output via pipes.  This amplifies and extends the utility of existing sensing and actuation approaches.  In addition to describing the design space of pipes, we also presented \systemnamenospace, a design tool for the creation of a series of tubes \cite{series-of-tubes} inside arbitrary 3D models, and discussed many techniques that combine nicely with pipes.  We fabricated several example objects using our tool to explore new points in the design space.  We look forward to extending our software contribution into future multi-material 3D modelling tools.